---
title: "LWT Data Exploration"
output:
  pdf_document: default
  html_document: default
---

install.packages("curl")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("gplots")
install.packages("corrplot")
install.packages("ggplot2")

```{r}
library(curl)
library(dplyr)
library(tidyverse)
library(gplots)
library(corrplot)
library(ggplot2)
```

## Just playing with my full dataset
```{r}
vervets_original<- curl("https://raw.githubusercontent.com/langley1/Vervet-Data/master/Vervet%20Behavior.csv")
vervets_original<- read.csv(vervets_original, header = TRUE)
head(vervets_original)
nrow(vervets_original)
summary(vervets_original)
names(vervets_original)
unique(vervets_original$Location)
unique(vervets_original$Age)
sort(table(vervets_original$Behavior), decreasing = TRUE)
```

## Importing Updated Dataset with edits given by Olivia
```{r}
vervets_full<- curl("https://raw.githubusercontent.com/langley1/Vervet-Data/master/LWT%20Vervet%20Dataset_Updated.csv")
vervets_full<- read.csv(vervets_full, header = TRUE)
head(vervets_full)
str(vervets_full)
summary(vervets_full)
```

## Creating unique IDs for each scan sample 
```{r}
vervets_full<- vervets_full %>% unite("Scan.info", Day, Month, Scan.time, sep = "_", remove = FALSE)  #this unites the 3 columns, (day, month, scan time) into one column called "Scan.info" so that I can create an ID based on each unique scan 
vervets_full<- transform(vervets_full, ScanID= as.numeric(factor(Scan.info))) #this creates a new column called "ScanID" and assigns its #s based on the Scan.info column 

tail(levels(factor(vervets_full$ScanID))) #There are a total of 1168 scans
```

## Breaking up the food into different categories and creating dfs for each
```{r}
food_HF<- vervets_full %>% filter(Descriptor %in% c("HF", "HFn", "HFg", "HFc", "Cr")) #human food and crops
food_SE<- vervets_full %>% filter(Descriptor %in% c("Sep", "Sed")) #Spilled enclosure food
food_Nat<- vervets_full %>% filter(Descriptor %in% c("N", "RV", "R")) #Natural and river
food_Water<- vervets_full %>% filter(Descriptor %in% c("ST", "H", "WT")) #sewage, hose, water

#Creating a new column called Descriptor Cateogory for the food groupings
food_HF$Descriptor.Cat<- "HF"
food_SE$Descriptor.Cat<- "SE"
food_Nat$Descriptor.Cat<- "NRV"
food_Water$Descriptor.Cat<- "W"

#Combing all of my dfs into one so that I can plot it by the Descript.Cat. Column
food_desc_ALL<- rbind(food_HF, food_SE, food_Nat, food_Water)
unique(food_desc_ALL$Descriptor.Cat)
```

## Plotting the food groupings 

### *by Age*
```{r}
food_age_plot <- ggplot(food_desc_ALL, aes(Descriptor.Cat, group = Age)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~Age)
food_age_plot
```

### *by Sex*
```{r}
food_sex_plot <- ggplot(food_desc_ALL, aes(Descriptor.Cat, group = Sex)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~Sex)
food_sex_plot
```

## Breaking up the locations into High/Medium/Low Human Interactions (as recommended by Olivia) and adding a new column
```{r}
high_human<- vervets_full %>% filter(Location %in% c("HO", "PG", "RO", "RS"))
med_human<- vervets_full %>% filter(Location %in% c("CG", "CP", "GP", "Ln", "PW", "WG"))
low_human<- vervets_full %>% filter(Location %in% c("F", "RF", "RB", "RV", "Oeb", "Oed", "Oep", "Ied"))

high_human$HumanLevel<- "H"
med_human$HumanLevel<- "M"
low_human$HumanLevel<- "L"

nrow(high_human) #1818
nrow(med_human) #5199
nrow(low_human) #3499
```

## Counting up the number of different behaviors in each group (high/med/low) in descending order
```{r}
high_human_behaviorcount<- high_human %>% 
  group_by(Behavior) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
nrow(high_human_behaviorcount) #13 behaviors (should Ag+/-, Th+/-, G+/-, etc. be counted as two different behaviors or just one?)

high_human_behaviorcount<- high_human %>%
  group_by(Behavior, ScanID) %>%
  summarize(count=n()) %>% 
  arrange(desc(count))

med_human_behaviorcount<- med_human %>% 
  group_by(Behavior) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
nrow(med_human_behaviorcount) #14 behaviors

low_human_behaviorcount<- low_human %>% 
  group_by(Behavior) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
nrow(low_human_behaviorcount) #13 behaviors

#Overall, there are roughly the same total number of unique behaviors seen across the three location categories- human level doesn't seem to affect behavioral diversity
```

## Finding the proportion of Adult Female behavior occurences in **high human** category

### *Total*
```{r}
AF_high_human<- high_human %>% filter(Age == "A", Sex == "F")
head(AF_high_human) #408 rows
```

### *Feeding/foraging*
```{r}
AF_Fe_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
head(AF_Fe_high_human) #179 rows

prop_AF_Fe_high_human<- (nrow(AF_Fe_high_human)/nrow(AF_high_human))*100
prop_AF_Fe_high_human #43.87
```

### *Grooming*
```{r}
AF_G_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
head(AF_G_high_human) #14 rows

prop_AF_G_high_human<- (nrow(AF_G_high_human)/nrow(AF_high_human))*100
prop_AF_G_high_human #8.82
```

### *Aggression*
```{r}
AF_Ag_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
head(AF_Ag_high_human) #3 rows

prop_AF_Ag_high_human<- (nrow(AF_Ag_high_human)/nrow(AF_high_human))*100
prop_AF_Ag_high_human #0.735
```

### *Predator Avoidance*
```{r}
AF_PA_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("PA"))
head(AF_PA_high_human)

prop_AF_PA_high_human<- (nrow(AF_PA_high_human)/nrow(AF_high_human))*100
prop_AF_PA_high_human #0.490
```

### *Self Groom*
```{r}
AF_SG_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("SG"))
head(AF_SG_high_human)

prop_AF_SG_high_human<- (nrow(AF_SG_high_human)/nrow(AF_high_human))*100
prop_AF_SG_high_human #0
```

### *Threatening*
```{r}
AF_Th_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Th+", "Th-"))
head(AF_Th_high_human)

prop_AF_Th_high_human<- (nrow(AF_Th_high_human)/nrow(AF_high_human))*100
prop_AF_Th_high_human #0
```

## Finding the proportion of Adult Female behavior occurences in **medium human** category

### *Total*
```{r}
AF_med_human<- med_human %>% filter(Age == "A", Sex == "F")
head(AF_med_human) #1,073 rows
```

### *Feeding/foraging*
```{r}
AF_Fe_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
head(AF_Fe_med_human) #467

prop_AF_Fe_med_human<- (nrow(AF_Fe_med_human)/nrow(AF_med_human))*100
prop_AF_Fe_med_human #43.52
```

### *Grooming*
```{r}
AF_G_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
head(AF_G_med_human) #60

prop_AF_G_med_human<- (nrow(AF_G_med_human)/nrow(AF_med_human))*100
prop_AF_G_med_human #5.59
```

### *Agression*
```{r}
AF_Ag_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
head(AF_Ag_med_human) #3

prop_AF_Ag_med_human<- (nrow(AF_Ag_med_human)/nrow(AF_med_human))*100
prop_AF_Ag_med_human #0.28
```

### *Predator Avoidance*
```{r}
AF_PA_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("PA"))
head(AF_PA_med_human)

prop_AF_PA_med_human<- (nrow(AF_PA_med_human)/nrow(AF_med_human))*100
prop_AF_PA_med_human #0.0
```

### *Self Groom*
```{r}
AF_SG_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("SG"))
head(AF_SG_med_human)

prop_AF_SG_med_human<- (nrow(AF_SG_med_human)/nrow(AF_med_human))*100
prop_AF_SG_med_human #0.0
```

### *Threatening*
```{r}
AF_Th_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Th+", "Th-"))
head(AF_Th_med_human)

prop_AF_Th_med_human<- (nrow(AF_Th_med_human)/nrow(AF_med_human))*100
prop_AF_Th_med_human #0.186
```


## Finding the proportion of Adult Female behavior occurences in **low human** category

### *Total*
```{r}
AF_low_human<- low_human %>% filter(Age == "A", Sex == "F")
head(AF_low_human) #786 rows
```

### *Feeding/foraging*
```{r}
AF_Fe_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
head(AF_Fe_low_human) #150

prop_AF_Fe_low_human<- (nrow(AF_Fe_low_human)/nrow(AF_low_human))*100
prop_AF_Fe_low_human #19.08
```

### *Grooming*
```{r}
AF_G_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
head(AF_G_low_human) #97

prop_AF_G_low_human<- (nrow(AF_G_low_human)/nrow(AF_low_human))*100
prop_AF_G_low_human #12.34
```

### *Agression*
```{r}
AF_Ag_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
head(AF_Ag_low_human) #8

prop_AF_Ag_low_human<- (nrow(AF_Ag_low_human)/nrow(AF_low_human))*100
prop_AF_Ag_low_human #1.02
```

### *Predator Avoidance*
```{r}
AF_PA_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("PA"))
head(AF_PA_low_human)

prop_AF_PA_low_human<- (nrow(AF_PA_low_human)/nrow(AF_low_human))*100
prop_AF_PA_low_human #0
```

### *Self Groom*
```{r}
AF_SG_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("SG"))
head(AF_SG_low_human)

prop_AF_SG_low_human<- (nrow(AF_SG_low_human)/nrow(AF_low_human))*100
prop_AF_SG_low_human #0
```

### *Threatening*
```{r}
AF_Th_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Th+", "Th-"))
head(AF_Th_low_human)

prop_AF_Th_low_human<- (nrow(AF_Th_low_human)/nrow(AF_low_human))*100
prop_AF_Th_low_human #0.127
```

## Creating a matrix with the 3 location categories and AF feeding, grooming, aggression, predator avoidance, self-grooming, and threatening proportions
```{r}
#Note: I removed the self-groom behavior from this matrix because the value was 0 for high, med and low category, no self-grooming occurred with adult females

AF_behaviors_table2<- matrix(c(prop_AF_Fe_high_human, prop_AF_Fe_med_human, prop_AF_Fe_low_human, prop_AF_G_high_human, prop_AF_G_med_human, prop_AF_G_low_human, prop_AF_Ag_high_human, prop_AF_Ag_med_human, prop_AF_Ag_low_human, prop_AF_PA_high_human, prop_AF_PA_med_human, prop_AF_PA_low_human, prop_AF_Th_high_human, prop_AF_Th_med_human, prop_AF_Th_low_human), ncol= 3, byrow = TRUE)
colnames(AF_behaviors_table2)<- c("High Human", "Med Human", "Low Human")
rownames(AF_behaviors_table2)<- c("Feeding", "Grooming", "Aggression", "Predator Avoidance", "Threatening")
AF_behaviors_table2<- as.table(AF_behaviors_table2)
AF_behaviors_table2

balloonplot(t(AF_behaviors_table2), main ="Adult Female Behaviors", xlab ="", ylab="",
            label = FALSE, show.margins = FALSE)
```

## Running a chi-square test for the high/med/low AF behaviors
```{r}
AF_chisq2 <- chisq.test(AF_behaviors_table2)
AF_chisq2

AF_chisq2$observed
round(AF_chisq2$expected,2)

round(AF_chisq2$residuals, 3) #This shows us our Pearson residuals (Cells with the highest absolute standardized residuals contribute the most to the total Chi-square score)
corrplot(AF_chisq2$residuals, is.cor = FALSE) #Allows us to visualize the residuals, blue is positive correlation and red is negative

contrib2 <- 100*AF_chisq2$residuals^2/AF_chisq2$statistic #The contribution (in %) of a given cell to the total Chi-square score
round(contrib2, 3)

corrplot(contrib2, is.cor = FALSE) #Visualizing the contribution %s
```

## Combining all of my human level dfs into one
```{r}
humanlevel_ALL<- rbind(high_human, med_human, low_human)
nrow(humanlevel_ALL) #10516
nrow(vervets_full) #10517 (header?)
unique(humanlevel_ALL$HumanLevel)
```

## Using full_join to combine my human level categories with the food descriptor categories I made... then I make another dataset which removes the NAs from the Food Descriptor column
```{r}
fulljoinDF<- full_join(humanlevel_ALL, food_desc_ALL, by = c("Day", "Month", "Poop.collected.y.n", "AM.PM", "Weather", "Loc.Desc", "Scan.time", "Age", "Sex", "Strata", "Behavior", "Descriptor", "Location"))
View(fulljoinDF)

fulljoinDF_noNA<- drop_na(fulljoinDF)
nrow(fulljoinDF_noNA)
```

## Plotting the food descriptor categories across the three human levels 
```{r}
ggplot(data = na.omit(fulljoinDF),aes(x= HumanLevel, fill=Descriptor.Cat, na.rm = TRUE)) + 
   geom_bar(position="fill") #There are a lot of NAs in the Descriptor Cat column so this removes them

#This plot shows that there are the most HF, "Human Foods" in the high and medium human level locations and spilled enclosure food only in the low level group
```

## Plotting the different age groups across the three human levels, grouped by age
```{r}
HumanLevel_Age_plot <- ggplot(humanlevel_ALL, aes(HumanLevel, group = Age)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~Age)
HumanLevel_Age_plot

# This plot shows that the percentage of adult, juvenile, and infant monkeys across the 3 human levels is pretty consistent. Overall, there are the most individuals in the medium human level group and the least in the high human level group. 
```

## Plotting the ages/sexs in the three human levels, filled bar format

### *By Age*
```{r}
ggplot(data = humanlevel_ALL, aes(x= HumanLevel, fill=Age)) + 
   geom_bar(position="fill")
```

### *By Sex*
```{r}
ggplot(data = humanlevel_ALL, aes(x= HumanLevel, fill=Sex)) + 
   geom_bar(position="fill") #Lots of unknowns
```

## Looking at ONLY feeding and foraging behaviors in the High Human category across different scan time chunks (i.e. 6-10am; 9-12pm; 12-15pm; 15-18pm)
```{r}
library(lubridate)
HighHum_Fe_DF<- humanlevel_ALL %>% filter(HumanLevel == "H", Behavior %in% c("Fe", "Fo")) #Filtering out only the Fe/Fo behaviors in the High Human location category 

HighHum_Fe_DF$Scan.time <- strptime(HighHum_Fe_DF$Scan.time, "%H:%M") #Must use strptime() first b/c this tells the 'strptime' function what format is currently in use, and exports a POSIX-compatible datetime.
HighHum_Fe_DF$Scan.time <- as.POSIXct(HighHum_Fe_DF$Scan.time) #Now you can use the as.POSIXct format; NOTE: this puts in today's current Year-Month-Day into the Scan.time column, I'm not sure how to remove this yet so I'm just going to keep it in for now

HighHum_Fe_6thr9<- with(HighHum_Fe_DF, HighHum_Fe_DF[hour(Scan.time) >= 6 & hour(Scan.time) < 9,]) #Now specifying which time chunk I'd like to look at; 6 through 10
head(HighHum_Fe_6thr9)

HighHum_Fe_9thr12<- with(HighHum_Fe_DF, HighHum_Fe_DF[hour(Scan.time) >= 9 & hour(Scan.time) < 12,]) 
head(HighHum_Fe_9thr12) #9 through 12

HighHum_Fe_12thr15<- with(HighHum_Fe_DF, HighHum_Fe_DF[hour(Scan.time) >= 12 & hour(Scan.time) < 15,]) 
head(HighHum_Fe_12thr15) #12 through 15

HighHum_Fe_15thr18<- with(HighHum_Fe_DF, HighHum_Fe_DF[hour(Scan.time) >= 15 & hour(Scan.time) < 18,]) 
head(HighHum_Fe_15thr18) #15 through 18
```

## Creating a new column in HighHum_Fe_DF called "Time Block"
```{r}
#First I have to create a new column called "time.block" in each of my dfs and specify what I want in that column
HighHum_Fe_6thr9$Time.Block<- "6-9"
HighHum_Fe_9thr12$Time.Block<- "9-12"
HighHum_Fe_12thr15$Time.Block<- "12-15"
HighHum_Fe_15thr18$Time.Block<- "15-18"

#Now I combine all of my dfs into one so that I can plot it by the Time.Block Column
HighHum_Fe_ALL<- rbind(HighHum_Fe_6thr9, HighHum_Fe_9thr12, HighHum_Fe_12thr15, HighHum_Fe_15thr18)
unique(HighHum_Fe_ALL$Time.Block)
```

## Plotting the number of feeding/foraging behaviors across the 4 time blocks
```{r}
TimeBlock_Fe_plot<- ggplot(data = HighHum_Fe_ALL, aes(x= Time.Block, fill=Behavior)) + 
   geom_bar((aes(fill = Behavior)))
TimeBlock_Fe_plot

#From this plot, we can see that there are the highest number of feeding and foraging behaviors during the 6-9am time block and the second highest during 12-15. Feeding and foraging occured the least during the 15-18 time block.
```
