---
title: "Malawi Vervet Data"
output: html_document
---

install.packages("curl")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("gplots")
install.packages("corrplot")
install.packages("ggplot2")

```{r}
library(curl)
library(dplyr)
library(tidyverse)
library(gplots)
library(corrplot)
library(ggplot2)
```

# Just playing with my full dataset
```{r}
vervets_original<- curl("https://raw.githubusercontent.com/langley1/Vervet-Data/master/Vervet%20Behavior.csv")
vervets_original<- read.csv(vervets_original, header = TRUE)
head(vervets_original)
tail(vervets_original)
nrow(vervets_original)
summary(vervets_original)
names(vervets_original)
unique(vervets_original$Location)
unique(vervets_original$Age)
sort(table(vervets_original$Behavior), decreasing = TRUE)
```

# Importing Updated DataSet 
```{r}
vervets_full<- curl("https://raw.githubusercontent.com/langley1/Vervet-Data/master/LWT%20Vervet%20Dataset_Updated.csv")
vervets_full<- read.csv(vervets_full, header = TRUE)
head(vervets_full)
str(vervets_full)
summary(vervets_full)
```

# Using the with() function to count the total number of individual scans conducted
```{r}
with(vervets_full, interaction(Day, Month, Scan.time, AM.PM, drop=TRUE)) #Using the full dataset because some of my location groupings overlap with a single scan/ need to include day and month so I get every unique scan

table_full<- aggregate(numeric(nrow(vervets_full)), vervets_full[c("Day", "Month", "Scan.time", "AM.PM")], length) #this just puts it in a nice table 
nrow(table_full) #1168 total scans
```

# Creating a subset of the data only looking at location that falls under "human habitation"
```{r}
human_habitation<- vervets_full %>% filter(Location %in% c("HCS", "HS", "HO", "RS", "CP", "PG", "GP")) 
nrow(human_habitation) #3370
unique(human_habitation$Scan.time) #142

i <- sapply(human_habitation, is.factor) #I'm turning this into characters to see if that helps with the boxplot attempts
human_habitation[i] <- lapply(human_habitation[i], as.character)
human_habitation %>% mutate_if(is.factor, as.character) -> human_habitation

str(human_habitation)

human_habitation$Behavior
HH_behav<- human_habitation$Behavior
str(HH_behav)
#boxplot(HH_behav, data= human_habitation) ##Non Numeric Error
str(human_habitation)
```

# Finding the proportion of Adult Female behavior occurences in **human habitation** category

*Total*
```{r}
AF_HH<- human_habitation %>% filter(Age == "A", Sex == "F") #Starting with total number of Adult Fem behaviors in HH
AF_HH #767
```

*Feeding/foraging*
```{r}
AF_Fe_HH<- human_habitation %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
AF_Fe_HH #443 rows

prop_AF_Fe_HH<- (nrow(AF_Fe_HH)/nrow(AF_HH))*100
prop_AF_Fe_HH #57.76
```

*Grooming*
```{r}
AF_G_HH<- human_habitation %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
AF_G_HH #40 rows

prop_AF_G_HH<- (nrow(AF_G_HH)/nrow(AF_HH))*100
prop_AF_G_HH #5.22
```

*Aggression*
```{r}
AF_Ag_HH<- human_habitation %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
AF_Ag_HH #3

prop_AF_Ag_HH<- (nrow(AF_Ag_HH)/nrow(AF_HH))*100
prop_AF_Ag_HH #0.39
```

# Creating dataframe for Forest/river location 
```{r}
Forest<- vervets_full %>% filter(Location %in% c("F", "RF", "RB", "RV", "WG"))
Forest

unique(Forest$Behavior) #24 total behaviors (more than HH)
unique(Forest$Scan.time) #142
```

# Using the with() function to count the total number of individual scans conducted among the Forest locations
```{r}
with(Forest, interaction(Day, Month, Scan.time, AM.PM, drop=TRUE))   # gives labels

table<- aggregate(numeric(nrow(Forest)), Forest[c("Day", "Month", "Scan.time", "AM.PM")], length) 
table
```

# Finding the proportion of Adult Female behavior occurences in **Forest** category

*Total*
```{r}
AF_F<- Forest %>% filter(Age == "A", Sex == "F") #includes all behaviors
AF_F #631
```

*Feeding/foraging*
```{r}
AF_Fe_F<- Forest %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
AF_Fe_F #97

prop_AF_Fe_F<- (nrow(AF_Fe_F)/nrow(AF_F))*100
prop_AF_Fe_F #15.37
```

*Grooming*
```{r}
AF_G_F<- Forest %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
AF_G_F #71 rows

prop_AF_G_F<- (nrow(AF_G_F)/nrow(AF_F))*100
prop_AF_G_F #11.25
```

*Agression*
```{r}
AF_Ag_F<- Forest %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
AF_Ag_F #5 rows

prop_AF_Ag_F<- (nrow(AF_Ag_F)/nrow(AF_F))*100
prop_AF_Ag_F #0.7923
```

# Creating dataframe for Outside/Inside Enclosures
```{r}
Out_enclosures<- vervets_full %>% filter(Location %in% c("Oep", "Oed", "Oeb"))
nrow(Out_enclosures) #212
unique(Out_enclosures$Scan.time) #142

In_enclosures<- vervets_full %>% filter(Location %in% c("Ied")) #No data included for the other inside enclosure locations (Iep and Ieb)
nrow(In_enclosures) #346
unique(In_enclosures$Scan.time) #142
```

# Finding the proportion of Adult Female behavior occurences in **outside enclosure** category

*Total*
```{r}
AF_OE<- Out_enclosures %>% filter(Age == "A", Sex == "F")
AF_OE #57
```

*Feeding/foraging*
```{r}
AF_Fe_OE<- Out_enclosures %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
AF_Fe_OE #2

prop_AF_Fe_OE<- (nrow(AF_Fe_OE)/nrow(AF_OE))*100
prop_AF_Fe_OE #3.51
```

*Grooming*
```{r}
AF_G_OE<- Out_enclosures %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
AF_G_OE #12

prop_AF_G_OE<- (nrow(AF_G_OE)/nrow(AF_OE))*100
prop_AF_G_OE #21.05
```

*Agression*
```{r}
AF_Ag_OE<- Out_enclosures %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
AF_Ag_OE #0

prop_AF_Ag_OE<- (nrow(AF_Ag_OE)/nrow(AF_OE))*100
prop_AF_Ag_OE #0
```

# Finding the proportion of Adult Female behavior occurences in **inside enclosure** category

*Total*
```{r}
AF_IE<- In_enclosures %>% filter(Age == "A", Sex == "F")
AF_IE #99
```

*Feeding/foraging*
```{r}
AF_Fe_IE<- In_enclosures %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
AF_Fe_IE #51

prop_AF_Fe_IE<- (nrow(AF_Fe_IE)/nrow(AF_IE))*100
prop_AF_Fe_IE #51.52
```

*Grooming*
```{r}
AF_G_IE<- In_enclosures %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
AF_G_IE #14

prop_AF_G_IE<- (nrow(AF_G_IE)/nrow(AF_IE))*100
prop_AF_G_IE #14.14
```

*Agression*
```{r}
AF_Ag_IE<- In_enclosures %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
AF_Ag_IE #3

prop_AF_Ag_IE<- (nrow(AF_Ag_IE)/nrow(AF_IE))*100
prop_AF_Ag_IE #3.03
```

# Creating a matrix with the 4 location categories and AF feeding/grooming/aggression proportions
```{r}
AF_behaviors_table<- matrix(c(prop_AF_Fe_F, prop_AF_Fe_HH, prop_AF_Fe_IE, prop_AF_Fe_OE, prop_AF_G_F, prop_AF_G_HH, prop_AF_G_IE, prop_AF_G_OE, prop_AF_Ag_F, prop_AF_Ag_HH, prop_AF_Ag_IE, prop_AF_Ag_OE), ncol= 4, byrow = TRUE)
colnames(AF_behaviors_table)<- c("Forest", "HH", "In Encl", "Out Encl")
rownames(AF_behaviors_table)<- c("Feeding", "Grooming", "Aggression")
AF_behaviors_table<- as.table(AF_behaviors_table)
AF_behaviors_table

balloonplot(t(AF_behaviors_table), main ="Adult Female Behaviors", xlab ="", ylab="",
            label = FALSE, show.margins = FALSE)
```

# Running a chi-square test
```{r}
AF_chisq <- chisq.test(AF_behaviors_table)
AF_chisq

AF_chisq$observed
round(AF_chisq$expected,2)

round(AF_chisq$residuals, 3) #This shows us our Pearson residuals (Cells with the highest absolute standardized residuals contribute the most to the total Chi-square score)
corrplot(AF_chisq$residuals, is.cor = FALSE) #Allows us to visualize the residuals, blue is positive correlationa and red is negative

contrib <- 100*AF_chisq$residuals^2/AF_chisq$statistic #The contribution (in %) of a given cell to the total Chi-square score
round(contrib, 3)

corrplot(contrib, is.cor = FALSE) #Visualizing the contribution %s
```


# Creating dataset for the remaining locations 
```{r}
Outside<- vervets_full %>% filter(Location %in% c("CG", "PW", "Ln", "RO"))
nrow(Out_enclosures)
```

# Looking at some of the individual behaviors and counting their occurences in human habituation subset
```{r}
unique(HH_behav) #There are 11 total behaviors

HH_F<- human_habitation %>% filter(Behavior %in% c("Fe", "Fo"))
HH_F$Loc_Cat<- "HH" #adding in a new column so that when I merge all of the Fe/Fo dfs, I can distinguish by location category
head(HH_F)
nrow(HH_F) #158 feeding or foraging bouts in HH
perc_HH_F<- (nrow(HH_F)/530)*100
perc_HH_F

HH_PL<- human_habitation %>% filter(Behavior == "PL")
nrow(HH_PL) #28 playing bouts in HH
perc_HH_PL<- (nrow(HH_PL)/530)*100
perc_HH_PL

HH_G<- human_habitation %>% filter(Behavior %in% c("G+", "G-", "Gr+", "Gr-"))
nrow(HH_G) #31 grooming bouts
perc_HH_G<- (nrow(HH_G)/530)*100
perc_HH_G
```

```{r}
x<- data.frame("Behavior Category" = c("Foraging/Feeding", "Playing", "Grooming"), "Percent" = c(perc_HH_F, perc_HH_PL, perc_HH_G))
x
```

# Looking at some individual behaviors for forest/river location
```{r}
Forest_F<- Forest %>% filter(Behavior %in% c("Fe", "Fo"))
nrow(Forest_F) #399 feeding/foraging bouts 
Forest_F$Loc_Cat<- "F" #adding in a new column so that when I merge all of the Fe/Fo dfs, I can distinguish by location category
head(Forest_F)

Forest_PL<- Forest %>% filter(Behavior == "PL")
nrow(Forest_PL) #295 playing bouts in Forest

Forest_G<- Forest %>% filter(Behavior %in% c("G+", "G-", "Gr+", "Gr-"))
nrow(Forest_G) #156 
```

# Pulling out Feeding/foraging behaviors from the In/Out enclosure categories
```{r}
OE_F<- Out_enclosures %>% filter(Behavior %in% c("Fe", "Fo"))
OE_F$Loc_Cat<- "OE" #adding in a new column so that when I merge all of the Fe/Fo dfs, I can distinguish by location category
head(OE_F)

IE_F<- In_enclosures %>% filter(Behavior %in% c("Fe", "Fo"))
IE_F$Loc_Cat<- "IE"
head(IE_F)
```

# Combinging all of my foraging/feeding dfs from each location category togeher so that I can make a bar chart and group by location (or sex, etc.)
```{r}
Fe_ALL<- rbind(HH_F, Forest_F, OE_F, IE_F)
unique(Fe_ALL$Loc_Cat)
```

# Creating a plot for feeding/foraging behaviors
```{r}
ggplot(vervets_full) +
  aes(x = Age, fill = factor(Behavior)) +
  geom_bar(position = "fill")

myplot <- ggplot(Fe_ALL, aes(Loc_Cat, group = Age)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~Age)
myplot
```

```{r}
unique(vervets_full$Descriptor)
unique(human_habitation$Descriptor)
human_habitation
```

```{r}
head(starwars)
starwars.g<- group_by(starwars,species)
starwars.g
summarize(starwars.g, avg= mean(height,na.rm = TRUE))
```

# Breaking up the food into different categories and creating dfs for each
```{r}
food_HF<- vervets_full %>% filter(Descriptor %in% c("HF", "HFn", "HFg", "HFc", "Cr"))
food_SE<- vervets_full %>% filter(Descriptor %in% c("Sep", "Sed"))
food_Nat<- vervets_full %>% filter(Descriptor %in% c("N", "RV", "R"))
food_Water<- vervets_full %>% filter(Descriptor %in% c("ST", "H", "WT"))

#Creating a new column called Descriptor Cateogory for the food groupings
food_HF$Descriptor.Cat<- "HF"
food_SE$Descriptor.Cat<- "SE"
food_Nat$Descriptor.Cat<- "NRV"
food_Water$Descriptor.Cat<- "W"

#Combing all of my dfs into one so that I can plot it by the Descript.Cat. Column
food_desc_ALL<- rbind(food_HF, food_SE, food_Nat, food_Water)
unique(food_desc_ALL$Descriptor.Cat)
```

# Plotting the food groupings by Age
```{r}
myplot2 <- ggplot(food_desc_ALL, aes(Descriptor.Cat, group = Age)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~Age)
myplot2
```

# Plotting the food groupings by Sex
```{r}
myplot3 <- ggplot(food_desc_ALL, aes(Descriptor.Cat, group = Sex)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~Sex)
myplot3
```

# Breaking up the locations into High/Medium/Low Human Interactions and adding a new column
```{r}
high_human<- vervets_full %>% filter(Location %in% c("HO", "PG", "RO", "RS"))
med_human<- vervets_full %>% filter(Location %in% c("CG", "CP", "GP", "Ln", "PW", "WG"))
low_human<- vervets_full %>% filter(Location %in% c("F", "RF", "RB", "RV", "Oeb", "Oed", "Oep", "Ied"))

high_human$HumanLevel<- "H"
med_human$HumanLevel<- "M"
low_human$HumanLevel<- "L"

nrow(high_human) #1816 - 1818
nrow(med_human) #5199 - 5199
nrow(low_human) #3484 - 3499
```

# Counting up the number of different behaviors in each group (high/med/low) in descending order
```{r}
high_human_behaviorcount<- high_human %>% 
  group_by(Behavior) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
nrow(high_human_behaviorcount) #13 behaviors (should Ag+/-, Th+/-, G+/-, etc. be counted as two different behaviors or just one?)

med_human_behaviorcount<- med_human %>% 
  group_by(Behavior) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
nrow(med_human_behaviorcount) #14 behaviors

low_human_behaviorcount<- low_human %>% 
  group_by(Behavior) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
nrow(low_human_behaviorcount) #13 behaviors
```

# Finding the proportion of Adult Female behavior occurences in **high human** category

*Total*
```{r}
AF_high_human<- high_human %>% filter(Age == "A", Sex == "F")
AF_high_human #408
```

*Feeding/foraging*
```{r}
AF_Fe_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
AF_Fe_high_human #179

prop_AF_Fe_high_human<- (nrow(AF_Fe_high_human)/nrow(AF_high_human))*100
prop_AF_Fe_high_human #43.87
```

*Grooming*
```{r}
AF_G_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
AF_G_high_human #14

prop_AF_G_high_human<- (nrow(AF_G_high_human)/nrow(AF_high_human))*100
prop_AF_G_high_human #8.82
```

*Agression*
```{r}
AF_Ag_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
AF_Ag_high_human #3

prop_AF_Ag_high_human<- (nrow(AF_Ag_high_human)/nrow(AF_high_human))*100
prop_AF_Ag_high_human #0.735
```

*Predator Avoidance*
```{r}
AF_PA_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("PA"))
AF_PA_high_human

prop_AF_PA_high_human<- (nrow(AF_PA_high_human)/nrow(AF_high_human))*100
prop_AF_PA_high_human #0.490
```

*Self Groom*
```{r}
AF_SG_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("SG"))
AF_SG_high_human

prop_AF_SG_high_human<- (nrow(AF_SG_high_human)/nrow(AF_high_human))*100
prop_AF_SG_high_human #0
```

*Threatening*
```{r}
AF_Th_high_human<- high_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Th+", "Th-"))
AF_Th_high_human

prop_AF_Th_high_human<- (nrow(AF_Th_high_human)/nrow(AF_high_human))*100
prop_AF_Th_high_human #0
```

# Finding the proportion of Adult Female behavior occurences in **medium human** category

*Total*
```{r}
AF_med_human<- med_human %>% filter(Age == "A", Sex == "F")
AF_med_human #1,073
```

*Feeding/foraging*
```{r}
AF_Fe_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
AF_Fe_med_human #467

prop_AF_Fe_med_human<- (nrow(AF_Fe_med_human)/nrow(AF_med_human))*100
prop_AF_Fe_med_human #43.52
```

*Grooming*
```{r}
AF_G_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
AF_G_med_human #60

prop_AF_G_med_human<- (nrow(AF_G_med_human)/nrow(AF_med_human))*100
prop_AF_G_med_human #5.59
```

*Agression*
```{r}
AF_Ag_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
AF_Ag_med_human #3

prop_AF_Ag_med_human<- (nrow(AF_Ag_med_human)/nrow(AF_med_human))*100
prop_AF_Ag_med_human #0.28
```

*Predator Avoidance*
```{r}
AF_PA_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("PA"))
AF_PA_med_human

prop_AF_PA_med_human<- (nrow(AF_PA_med_human)/nrow(AF_med_human))*100
prop_AF_PA_med_human #0.0
```

*Self Groom*
```{r}
AF_SG_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("SG"))
AF_SG_med_human

prop_AF_SG_med_human<- (nrow(AF_SG_med_human)/nrow(AF_med_human))*100
prop_AF_SG_med_human #0.0
```

*Threatening*
```{r}
AF_Th_med_human<- med_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Th+", "Th-"))
AF_Th_med_human

prop_AF_Th_med_human<- (nrow(AF_Th_med_human)/nrow(AF_med_human))*100
prop_AF_Th_med_human #0.186
```


# Finding the proportion of Adult Female behavior occurences in **low human** category

*Total*
```{r}
AF_low_human<- low_human %>% filter(Age == "A", Sex == "F")
AF_low_human #786
```

*Feeding/foraging*
```{r}
AF_Fe_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Fe", "Fo"))
AF_Fe_low_human #150

prop_AF_Fe_low_human<- (nrow(AF_Fe_low_human)/nrow(AF_low_human))*100
prop_AF_Fe_low_human #19.08
```

*Grooming*
```{r}
AF_G_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("G+", "G-"))
AF_G_low_human #97

prop_AF_G_low_human<- (nrow(AF_G_low_human)/nrow(AF_low_human))*100
prop_AF_G_low_human #12.34
```

*Agression*
```{r}
AF_Ag_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Ag+", "Ag-"))
AF_Ag_low_human #8

prop_AF_Ag_low_human<- (nrow(AF_Ag_low_human)/nrow(AF_low_human))*100
prop_AF_Ag_low_human #1.02
```

*Predator Avoidance*
```{r}
AF_PA_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("PA"))
AF_PA_low_human

prop_AF_PA_low_human<- (nrow(AF_PA_low_human)/nrow(AF_low_human))*100
prop_AF_PA_low_human #0
```

*Self Groom*
```{r}
AF_SG_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("SG"))
AF_SG_low_human

prop_AF_SG_low_human<- (nrow(AF_SG_low_human)/nrow(AF_low_human))*100
prop_AF_SG_low_human #0
```

*Threatening*
```{r}
AF_Th_low_human<- low_human %>% filter(Age == "A", Sex == "F", Behavior %in% c("Th+", "Th-"))
AF_Th_low_human

prop_AF_Th_low_human<- (nrow(AF_Th_low_human)/nrow(AF_low_human))*100
prop_AF_Th_low_human #0.127
```

# Creating a matrix with the 4 location categories and AF feeding, grooming, aggression, predator avoidance, self-grooming, and threatening proportions
```{r}
AF_behaviors_table2<- matrix(c(prop_AF_Fe_high_human, prop_AF_Fe_med_human, prop_AF_Fe_low_human, prop_AF_G_high_human, prop_AF_G_med_human, prop_AF_G_low_human, prop_AF_Ag_high_human, prop_AF_Ag_med_human, prop_AF_Ag_low_human, prop_AF_PA_high_human, prop_AF_PA_med_human, prop_AF_PA_low_human, prop_AF_SG_high_human, prop_AF_SG_med_human, prop_AF_SG_low_human, prop_AF_Th_high_human, prop_AF_Th_med_human, prop_AF_Th_low_human), ncol= 3, byrow = TRUE)
colnames(AF_behaviors_table2)<- c("High Human", "Med Human", "Low Human")
rownames(AF_behaviors_table2)<- c("Feeding", "Grooming", "Aggression", "Predator Avoidance", "Self-Grooming", "Threatening")
AF_behaviors_table2<- as.table(AF_behaviors_table2)
AF_behaviors_table2

balloonplot(t(AF_behaviors_table2), main ="Adult Female Behaviors", xlab ="", ylab="",
            label = FALSE, show.margins = FALSE)
```

# Running a chi-square test for the high/med/low AF behaviors
```{r}
AF_chisq2 <- chisq.test(AF_behaviors_table2)
AF_chisq2

AF_chisq2$observed
round(AF_chisq2$expected,2)

round(AF_chisq2$residuals, 3) #This shows us our Pearson residuals (Cells with the highest absolute standardized residuals contribute the most to the total Chi-square score)
corrplot(AF_chisq2$residuals, is.cor = FALSE) #Allows us to visualize the residuals, blue is positive correlationa and red is negative

contrib2 <- 100*AF_chisq2$residuals^2/AF_chisq2$statistic #The contribution (in %) of a given cell to the total Chi-square score
round(contrib2, 3)

corrplot(contrib2, is.cor = FALSE) #Visualizing the contribution %s
```

#Combining all of my human level dfs into one
```{r}
humanlevel_ALL<- rbind(high_human, med_human, low_human)
nrow(humanlevel_ALL) #10516
nrow(vervets_full) #10517 (header?)
unique(humanlevel_ALL$HumanLevel)
```

```{r}
nrow(humanlevel_ALL) #10516
nrow(food_desc_ALL) #3298

fulljoinDF<- full_join(humanlevel_ALL, food_desc_ALL, by = c("Day", "Month", "Poop.collected.y.n", "AM.PM", "Weather", "Loc.Desc", "Scan.time", "Age", "Sex", "Strata", "Behavior", "Descriptor", "Location"))
View(fulljoinDF)
```

# Plotting the food descriptor categories across the three human levels 
```{r}
ggplot(data = na.omit(fulljoinDF),aes(x= HumanLevel, fill=Descriptor.Cat, na.rm = TRUE)) + 
   geom_bar(position="fill") #There are a lot of NAs in the Descriptor Cat column so this removes them
```

# Plotting the different age groups across the three human levels, grouped by age
```{r}
myplot3 <- ggplot(humanlevel_ALL, aes(HumanLevel, group = Age)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~Age)
myplot3
```

# Plotting the ages/sexs in the three human levels, filled bar format
```{r}
ggplot(data = fulljoinDF, aes(x= HumanLevel, fill=Age)) + 
   geom_bar(position="fill")

ggplot(data = fulljoinDF, aes(x= HumanLevel, fill=Sex)) + 
   geom_bar(position="fill") #Lots of unknowns
```


